---
title: "ESM 244 Lab 1"
author: "Shuying Yu"
date: "1/8/2022"
output: html_document
---

```{r setup, include = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(sf)
library(tmap)

#Updates all packages
### update.packages(ask = FALSE)

#Use control+shift+enter anywhere in code chunk to run eveyrthing inside chunk
```


# Read in the data

```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))
```


```{r}
summary(sf_trees)
```


# Part 1: wrangling and ggplot review

*Example 1:* Find counts observation by `legal_status` and wrangle a bit

```{r}
#Method 1: group_by %>% summarize()

sf_trees %>% 
  group_by(legal_status) %>% 
  summarize(tree_count = n())



#Method 2: different way plus a few functions
top_5_status <- sf_trees %>% 
  
  #Built in to count things you want to group by
  count(legal_status) %>% 
  
  #drop rows with NA in legal status column
  drop_na(legal_status) %>% 
  
  #rename n to tree_count
  rename(tree_count = n) %>% 
  
  #relocate column to be first
  relocate(tree_count) %>% 
  
  #top 5 tree_count
  slice_max(tree_count, n = 5) %>% 
  
  #arrange by tree_count highest to lowest
  #or arange(desc(tree_count))
  arrange(-tree_count)
```


Make a graph of the top 5

```{r}
#Alphabetical order
ggplot(data = top_5_status,
       aes(x = legal_status,
           y = tree_count)) +
  
  geom_col()



ggplot(data = top_5_status,
       
       #reorder legal status by tree count
       aes(x = fct_reorder(legal_status, tree_count),
           y = tree_count)) +
  
  geom_col(fill = "darkgreen") +
  
  #Change labels
  labs(x = "Lagel Status",
       y = "Tree Count") +
  
  #Rotate coordinate
  coord_flip() +
  
  #Change theme
  theme_minimal()
```

*Example 2:* Only keep observations where legal status is "Permitted Site" and caretaker is "MTA", and store as `permitted_data_df`


```{r}
#Check unique legal status labels
sf_trees$legal_status %>%  unique()
unique(sf_trees$caretaker)

#Subset and filter
permitted_data_df <- sf_trees %>% 
  filter(legal_status == "Permitted Site",
         caretaker == "MTA")
```

*Example 3:* Only keep Acaia trees, then only keep columns `legal_status`, `date`, `latitude`, and store as `blackwood_acaia_df`

```{r}
blackwood_acaia_df <- sf_trees %>% 
  
  #in species, detect rows where these words are mentioned or used 
  filter(str_detect(species, "Blackwood Acacia")) %>% 
  
  #rename latitide and longitude
  select(legal_status, date,
         lat = latitude,
         lon = longitude)


#Make a little graph of locations
ggplot(data = blackwood_acaia_df,
       aes(x = lon, y = lat)) +
  geom_point(color = "darkgreen")
```

*Example 4:* use `tidyr::separate`

Speparate sepcies name with "::"

```{r}
sf_trees_sep <- sf_trees %>% 
  
  #separate species column
  #name new columns as
  #separator as ::
  separate(species, into = c("species_scientific", "species_common"),
           sep = "::")
```


*Example 5:* use `tidyr::unite`

```{r}
ex5 <- sf_trees %>% 
  
  #create new column id_status
  #combine tree_id and legal_status
  #sepeate with _C00L_
  unite("id_status",
        tree_id, legal_status,
        sep = "_C00L")
```

























